---
title: "Stochastically Perturbed Parmaters [SPP]"
author: "Andrew Singleton"
date: "Preliminary point verification for SOP 1"
output: 
  revealjs::revealjs_presentation:
    theme: league
    transition: fade
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE, dev.args = list(bg = "transparent"), collapse = TRUE,
  fig.height = 5
)

library(harp)
library(tidyverse)
library(here)
library(patchwork)
library(parallel)

plot_colours <- data.frame(
  mname = c("REF", "REF_new"),
  colour = c("#FFC080", "#C04040")
)

theme_alertness <- function (
  base_size = 11, base_family = "", base_line_size = base_size/22,
  base_rect_size = base_size/22
) {
  
  ggplot2::theme_grey(base_size = base_size, base_family = base_family,
    base_line_size = base_line_size, base_rect_size = base_rect_size) %+replace%
    ggplot2::theme(
      text = element_text(
        family = base_family,
        face = "plain",
        colour = "grey72",
        size = base_size,
        lineheight = 0.9,
        hjust = 0.5,
        vjust = 0.5,
        angle = 0,
        margin = margin(),
        debug = FALSE
      ),
      axis.text         = element_text(size = rel(0.8), colour = "grey50"),
      plot.background   = ggplot2::element_rect(fill = "#333333", colour = NA),
      panel.background  = ggplot2::element_rect(fill = "#333333", colour = NA),
      panel.border      = ggplot2::element_rect(fill = NA, colour = "grey92"),
      panel.grid        = ggplot2::element_line(colour = "grey30"),
      panel.grid.minor  = ggplot2::element_line(size = ggplot2:::rel(0.5)),
      strip.background  = ggplot2::element_rect(fill = "#4E5C68", colour = "grey92"),
      strip.text        = ggplot2::element_text(colour = "grey92"),
      legend.key        = ggplot2::element_rect(fill = "#333333", colour = NA),
      legend.background = ggplot2::element_rect(fill = "#333333", colour = NA),
      complete          = TRUE
    )
}

verif <- readRDS(here("data", "verification", "sop1_spp.rds"))
t2m <- readRDS(here("data", "verification", "t2m_timeSeries_sop1.rds"))
verif_t2m <- readRDS(here("data", "verification", "t2m_timeSeries_sop1_verif.rds"))

basic_plot <- function(vvv) {
  
  p1 <- plot_point_verif(
    vvv, spread_skill, plot_num_cases = FALSE, plot_caption = "", 
    colour_table = plot_colours, plot_title = "Spread : Skill - SOP 1"
  ) + 
    scale_linetype_manual(values = c(1, 3), labels = c("RMSE", "Spread")) +
    guides(linetype = guide_legend(NULL, override.aes = list(colour = "#CCCCCC"))) +
    labs(x = "Lead time [h]")
  
  p2 <- plot_point_verif(
    vvv, mean_bias, plot_num_cases = FALSE, plot_caption = "", 
    colour_table = plot_colours, plot_title = "Mean Bias - SOP 1"
  ) +
    labs(x = "Lead time [h]")
  
  p3 <- plot_point_verif(
    vvv, crps, plot_num_cases = FALSE, plot_caption = "", 
    colour_table = plot_colours, plot_title = "CRPS - SOP 1"
  ) +
    labs(x = "Lead time [h]")
  
  p4 <- plot_point_verif(
    vvv, normalized_rank_histogram, 
    colour_table = plot_colours, plot_title = "Rank Histogram - SOP 1"
  )
  
  (p1 + p2) / (p3 + p4) + plot_layout(guides = "collect") &
    theme_alertness(base_size = 12) &
    theme(legend.position = "bottom")
}

profile_plot <- function(vvv, score) {
  score <- enquo(score)
  vvv <- filter_list(vvv, leadtime %% 6 == 0) %>% 
    mutate_list(
      leadtime = fct_inorder(
        paste0("T + ", formatC(leadtime, width = 2, flag = "0"), "h")
      )
    )

  plot_profile_verif(
    vvv, !!score, facet_by = vars(leadtime), colour_table = plot_colours,
    plot_title = "Spread : Skill - SOP 1"
  ) + 
    scale_linetype_manual(values = c(1, 3), labels = c("RMSE", "Spread")) +
    guides(linetype = guide_legend(NULL, override.aes = list(colour = "#CCCCCC"))) +
    labs(x = "Lead time [h]") +
    theme_alertness()
}

```

<style>
  .reveal {font-family: comfortaa;}
  .reveal h1, .reveal h2, .reveal h3, .reveal h4, .reveal h5, .reveal h6 {text-transform:none; font-family:comfortaa;}
  .reveal section img { background:none; border:none; box-shadow:none; }
  .reveal table { border-width:0px; }
  .reveal table td { bordoer-width:0px; }
  .reveal pre code {background:#222;}
  .reveal pre {margin: 7px auto;}
  code span.kw {color:yellow}
  code span.dt {color: tomato}
</style>

## __What parameters do we perturb?__
- <span style="color:#FF7777">Cloud thickness threshold for shallow / deep convection</span>
- <span style="color:#BBBBFF">Cloud ice content impact on cloud thickness</span>
- <span style="color:#BBBBFF">Ice nucleii concentration</span>
- <span style="color:#7777FF">Saturation limit sensitivity for condensation*</span>
- <span style="color:#7777FF">Kogan autoconversion speed</span>
- <span style="color:#7777FF">Kogan sub grid scale (cloud fraction) sensitivity*</span>
- <span style="color:#DDDD55">Impact of graupel on radiation</span>
- <span style="color:#DDDD55">Impact of snow on radiation*</span>
- <span style="color:#99FF99">Top entrainment efficiency</span>
- <span style="color:#99FF99">Stable conditions length scale</span>
- <span style="color:#99FF99">Asymptotic free atmosphere length scale</span>

## ___However...___ __This Presentation is not as advertised!__

- Original reference runs had problems with warm perturbed members in winter and dry perturbed members in summer
- Original reference did not include modelling of snow on sea ice
- Some default physics parameter values are changed for SPP
  
---

- The reference is re-run with 
  - 6 perturbed members
  - No surface data assimilation for perturbed members
  - Snow on sea ice modelling
  - Physics parameter defaults changed 
    
## __Point verification for (near) surface parameters__

## __2m Temperature__
```{r t2m-plot, fig.width = 9, fig.height=6.5, fig.align="center"}
basic_plot(verif$T2m)
```

## __2m Temperature [members]__
```{r t2m-members}
ggplot(
  mutate(
    verif$T2m$det_summary_scores,
    type = case_when(member == "mbr000" ~ "Control", TRUE ~ "Perturbed")
  ),
  aes(leadtime, bias, group = member, colour = type)
) + 
  geom_line(size = 1) + 
  scale_colour_manual(NULL, values = c(Perturbed = "#A99080", Control = "#4040FF")) +
  facet_wrap(vars(mname)) + 
  geom_line(
    data = filter(verif$T2m$det_summary_scores, member == "mbr000"), 
    colour = "#4040FF", size = 1.2
  ) + 
  scale_x_continuous(breaks = seq(0, 180, 6)) +
  labs(x = "Lead Time [h]", y = bquote(Bias~"["*degree*C*"]")) +
  theme_alertness()
```

## __2m Temperature "first guess"__
```{r t2m-fg, fig.width = 9, fig.height=6.5, fig.align="center"}
p1 <- plot_point_verif(
  verif_t2m, mean_bias, filter_by = vars(leadtime == 3), x_axis = validdate,
  plot_num_cases = FALSE, colour_table = plot_colours, plot_caption = ""
) + 
  labs(x = "Date", y = bquote(Mean~Bias~"["*degree*C*"]"))

p2 <- ggplot(
  filter(t2m$REF, leadtime == 3),
  aes(validdate, T2m, group = validdate)
) + 
  geom_boxplot(colour = "#A99080", fill = "#A99080", alpha = 0.1) + 
  labs(
    x = "Date", 
    y = bquote("2m"~Temperature~"["*degree*C*"]"), 
    title = "Observed 2m Temperature"
  )

p1 / p2 & theme_alertness()
```

## __Temperature [upper air]__
```{r t-ss, fig.width = 9, fig.height=6.5, fig.align="center"}
profile_plot(verif$T, spread_skill)
```

## __2m Humidity__
```{r q2m-plot, fig.width = 9, fig.height=6.5, fig.align="center"}
basic_plot(verif$Q2m)
```

## __2m Humidity [members]__
```{r q2m-members}
ggplot(
  mutate(
    verif$Q2m$det_summary_scores,
    type = case_when(member == "mbr000" ~ "Control", TRUE ~ "Perturbed")
  ),
  aes(leadtime, bias, group = member, colour = type)
) + 
  geom_line(size = 1) + 
  scale_colour_manual(NULL, values = c(Perturbed = "#A99080", Control = "#4040FF")) +
  facet_wrap(vars(mname)) + 
  geom_line(
    data = filter(verif$Q2m$det_summary_scores, member == "mbr000"), 
    colour = "#4040FF", size = 1.2
  ) + 
  scale_x_continuous(breaks = seq(0, 180, 6)) +
  labs(x = "Lead Time [h]", y = bquote(Bias~"["*kg*"."*kg^-1*"]")) +
  theme_alertness()
```

## __Relative Humidity [upper air]__
```{r t-rh, fig.width = 9, fig.height=6.5, fig.align="center"}
profile_plot(verif$RH, spread_skill)
```

## __10m Wind Speed__
```{r s10m-plot, fig.width = 9, fig.height=6.5, fig.align="center"}
basic_plot(verif$S10m)
```

## __10m Wind Speed [members]__
```{r s10m-members}
ggplot(
  mutate(
    verif$S10m$det_summary_scores,
    type = case_when(member == "mbr000" ~ "Control", TRUE ~ "Perturbed")
  ),
  aes(leadtime, bias, group = member, colour = type)
) + 
  geom_line(size = 1) + 
  scale_colour_manual(NULL, values = c(Perturbed = "#A99080", Control = "#4040FF")) +
  facet_wrap(vars(mname)) + 
  geom_line(
    data = filter(verif$S10m$det_summary_scores, member == "mbr000"), 
    colour = "#4040FF", size = 1.2
  ) + 
  scale_x_continuous(breaks = seq(0, 180, 6)) +
  labs(x = "Lead Time [h]", y = bquote(Bias~"["*m*"."*s^-1*"]")) +
  theme_alertness()
```

## __Wind Speed [upper air]__
```{r s-ss, fig.width = 9, fig.height=6.5, fig.align="center"}
profile_plot(verif$S, spread_skill)
```

## __6h Precipitation__
```{r pcp6h-plot, fig.width = 9, fig.height=6.5, fig.align="center"}
basic_plot(filter_list(verif$AccPcp6h, num_cases > 10))
```

## __6h Precipitation [members]__
```{r pcp6-members}
ggplot(
  mutate(
    filter(verif$AccPcp6h$det_summary_scores, num_cases > 10),
    type = case_when(member == "mbr000" ~ "Control", TRUE ~ "Perturbed")
  ),
  aes(leadtime, bias, group = member, colour = type)
) + 
  geom_line(size = 1) + 
  scale_colour_manual(NULL, values = c(Perturbed = "#A99080", Control = "#4040FF")) +
  facet_wrap(vars(mname)) + 
  geom_line(
    data = filter(verif$AccPcp6h$det_summary_scores, member == "mbr000", num_cases > 10), 
    colour = "#4040FF", size = 1.2
  ) + 
  scale_x_continuous(breaks = seq(0, 180, 6)) +
  labs(x = "Lead Time [h]", y = bquote(Bias~"["*kg*"."*m^-2*"]")) +
  theme_alertness()
```

## __Total Cloud Cover__
```{r cctot-plot, fig.width = 9, fig.height=6.5, fig.align="center"}
basic_plot(verif$CCtot)
```

## __Total Cloud Cover [members]__
```{r cctot-members}
ggplot(
  mutate(
    verif$CCtot$det_summary_scores,
    type = case_when(member == "mbr000" ~ "Control", TRUE ~ "Perturbed")
  ),
  aes(leadtime, bias, group = member, colour = type)
) + 
  geom_line(size = 1) + 
  scale_colour_manual(NULL, values = c(Perturbed = "#A99080", Control = "#4040FF")) +
  facet_wrap(vars(mname)) + 
  geom_line(
    data = filter(verif$CCtot$det_summary_scores, member == "mbr000"), 
    colour = "#4040FF", size = 1.2
  ) + 
  scale_x_continuous(breaks = seq(0, 180, 6)) +
  labs(x = "Lead Time [h]", y = bquote(Bias~"[fraction]")) +
  theme_alertness()
```

## __Geopotential Height [upper air]__
```{r z-ss, fig.width = 9, fig.height=6.5, fig.align="center"}
profile_plot(verif$Z, spread_skill)
```

## __Summary__
- Warming of perturbed members reduced
- Night time wet bias removed in perturbed members
- Large positive bias in initial cloud cover for perturbed members removed
- Overall reduction in spread for all parameters
- Less spread throughout atmosphere

